Solutions to the potential for unintended conservation consequences of MPAs

### 2021-10-04 documents/03_marlin_paper.Rmd
testing reverse optimization

group_var <- "placement_strategy"

results <- placement_experiments %>%
  mutate(temp = raw_experiment_results) %>%
  mutate(state_id = map(raw_experiment_results, names)) %>%
  unnest(cols = c(temp, state_id)) %>%
  mutate(obj = map(temp, "obj"),
         state_id = as.numeric(state_id)) %>%
  unnest(cols = obj) %>%
  mutate(
    critter = fct_recode(
      critter,
      "Shark" = "carcharhinus longimanus",
      "Tuna" = "katsuwonus pelamis",
      "Marlin" = "kajikia audax"
    )
  )

if (any(results$econ[results$prop_mpa == 1] > 0)) {
  stop("something has gone wrong: yields present with 100% mpa")
}

# compare to BAU

results <- results %>%
  group_by(placement_id, state_id, critter) %>%
  mutate(bau_biodiv = biodiv[prop_mpa == 0],
         bau_econ = econ[prop_mpa == 0])

excols <-
  state_experiments %>%
  select(state_id, data) %>%
  unnest(cols = data) %>%
  select(state_id:sigma_hab) %>%
  ungroup() %>%
  unique()

total_results <- results %>%
  group_by(
    placement_id,
    placement_strategy,
    prop_critters_considered,
    placement_error,
    state_id,
    prop_mpa
  ) %>%
  summarise(
    loss = mean(biodiv < bau_biodiv),
    n_loss = sum(biodiv < bau_biodiv),
    signif_loss = mean(biodiv < (0.9 * bau_biodiv)),
    total_loss = sum(biodiv - bau_biodiv),
    econ = sum(econ),
    biodiv = sum(biodiv),
    bau_econ = sum(bau_econ),
    bau_biodiv = sum(bau_biodiv)
  ) %>%
  ungroup() %>%
  mutate(delta_biodiv = biodiv / bau_biodiv - 1,
         abs_delta_biodiv = biodiv - bau_biodiv,
         delta_econ = econ / bau_econ - 1) %>%
  left_join(excols, by = "state_id")
```

@@

@@

## Illustrative Experiments

@@

## WCPO Case Study

@@

## Illustrative Experiments

We focused first on the range of impacts that MPAs can have on mature
biomass across the range of variables explored. To illustrate, we pulled
out two example simulations, one where MPAs provide relatively large
benefits across all three species, and a second where MPAs actually
produced substantial losses in some species....

@@

Looking at the full
`r n_distinct(total_results$placement_id) * n_distinct(total_results$state_id)`
experiments run, the results were highly variable, with MPAs capable of
producing substantial conservation gains to substantial conservation
losses and everything in between. We do not focus into the individual
effects of particular variables as we found them to be highly dependent
on the exact question asked and the set of data included, and since we
cannot ascribe any actual likelihood to any individual state of nature
simulated here.

@@

Make clear that a goal of all this complexity was to see if there were
any clear emergent properties (this thing matters way more than this
thing). We highlight some of those, but given the complexity we found we
present the complexity of interacting ways in which different things can
effect the conservation outcomes of MPAs

XX after looking through these should widdle it down to only things you
actually examine, annoying / no point to having a bunch of unexplained
minor variation in the mix relatede to things like ontogenetic shifts)
xx

xx the problem here. These things all interact in so many complex ways
that it's hard to look at either graphically or statistically, and maybe
for the best. i.e. the answer isn't "if habitat overlap is less than XX%
you might have a problem, otherwise you're fine", but rather that a
bunch of things can interact to cause potential problems in complex
ways. but that's hard to explain xx.

Big message: even constrained by the same initial depletion and core
biology (e.g. movement rates), a wide range of outcomes are possible for
each species for a given MPA size, dependent on the design strategy,
habitat overlap, migrations, quality of information, etc. Across the
board as MPA size increases the ratio of positive to negative outcomes
increases, until all species achieve their maximum conservation
potential in every iteration at 100% MPA. HOwever, every species
experienced a set of simulations, roughyl xx%, that resulted in net
losses in SSB/SSB0 relative to BAU. Losses were smaller than potential
gaines, but reached meaningful levels in some species, and interestingly
peaked in many cases right before max MPA (reflecting concentration of
fishing in small amounts of habitat in habited by the species).Important
to stress in here that there are no probabilities in this, no way of
knowing which state is more probable than others, just given the range
of things simulated more ways for a bigger positive effect. Also make
clear though, biggish losses (10% or more) possible, but rare.

@@

To address these questions, we ran a more limited "case study" set of
simulations, based on the dynamics of the WCPO tuna fisheries. Rather
than iterating over different movement dynamics and habitat
configurations, we fix habitat layers on the values predicted by
@bradley2021, and xx other things we fix xx.

>>>>

@@

### Fleet Dynamics

The key feature of `marlin` for the purposes of this project is the
ability to capture technical interactions across species. While marlin
does not simulate species interactions, fishing fleets in marlin are
capable of targeting and affecting multiple species simultaneously. This
allows us to capture for example the potential impacts of displacement
of effort from one area into another across multiple species.

A fishing fleet or *metier* is defined in `marlin` by a total pool of
effort that can have different catchability, prices, and selectivity for
individual species within a simulation. For example, both a longline and
purse seine fleet may capture bigeye (*Thunnus obesus*) and skipjack
(*Katswamus pelamus*), but the longline fleet is much more likely to
catch bigeye than skipjack, and *vice versa*. The fishing fleet then
generates catches, revenues, and profits from individual species that it
targets, and makes decisions based on its total objective function
across all species.

Revenues *R* for fleet *f* in time step *t* are given by

$$
 R_{t,p,f} = \sum_s^S \sum_a^A \Pi_{s,f}\frac{\alpha_{a,f}q_{s,f}E_{t,p,f}}{F_{t,p,s,a}} \times Y_{t,p,s,a}
$$

where *p* is patch, *s* is species, *a* is age, $\alpha$ is selectivity
at age, *q* is catchability , *F* is total fishing mortality, and *Y*
are total catches.

$$
Y_{t,p,s,a} = \frac{F_{t,p,s,a}}{z_{t,p,s,a}} \times b_{t,p,s,a} \times (1 - e^{-z_{t,p,s,a}}) 
$$

where *F* is total fishing mortality, given by

$$
F_{t,p,s,a} = \sum_f^F\alpha_{a,f}q_{s,f}E_{t,p,f}
$$

and

$$
 z_{t,p,s,a} = F_{t,p,s,a} + m_{s,a}
$$

Where *m* is natural mortality.

This means that each species experiences a total mortality at age in a
time step in a patch, individual fractions of which are portioned off as
catches and subsequently revenues for each fleet. Given revenues, we
then calculate profits \$\\phi\$ per

$$
\phi_{t,p,f} = R_{t,p,f} - c_f(E_{t,p,f} + 1)_{t,p,f}
^{1.3}$$

where *c* is cost per unit effort and, the 1.3 exponent is in the manner
of @costello2016a . The addition of one to the total effort is to
constrain the value to be greater than 1, preventing undesired behavior
where fishing effort actually gets cheaper if total effort is less than
1.

The important thing to note here then is that the revenues and profits
for a given fleet are a function of the prices and catchability of
individual species, but that mortality itself is independent of price.
So, fleet that has selectivity and catchability greater than zero can
exert fishing mortality on a species even if it obtains no value from
that species.

The fleet then decides how to allocate its effort in space based on one
of four possible spatial allocation strategies. The ideal free
distribution (IFD) (citation) is the standard method for distributing
fishing fleets in spaces. However, the IFD presents a number of
complications for our study. First, in order to simplify the assessment
of the impacts of different fishing mortality rates we simulate
scenarios with a fixed total pool of effort. As MPAs change the spatial
distribution of profits though, the total amount of effort required to
satisfy the IFD may not equal the total amount of effort required by the
model. Second, even if we were to implement the IFD for a single fleet,
the model is capable of tracking multiple fleets at once. This means
that the IFD for a single fleet would be a game-theoretic outcome of the
actions of each of the individual fleet conditional on the actions of
all other fleets. While possible to implement this would slow down our
model runs to the point of practically preventing large-scale evaluation
of the effects of MPAs on bycatch.

As such we explored a series of "next best" fleet distribution
algorithms. While not the IFD in any individual time step, over time
they start to approximate the IFD (check this) as each assumes that
fleets base their decisions for the current time step on the outcomes in
the prior timestep, meaning that the impacts of the actions of other
fishing fleets are somewhat internalized. The four possible fleet
distribution algorithms are

1.  Revenue per unit effort (RPUE): The fleet distributes itself in
    space based on revenue per unit effort
2.  Revenue: The fleet distributes itself in space based on revenue
3.  Profit per unit effort (PPUE): The fleet distributes itself in space
    based on profit per unit effort
4.  Profit: The fleet distributes itself in space based on profit

### Population Model

@@

## Generalized Experiments

@@

## Case Study Application

@@

what we found was a series of neccesary but not sufficient conditions

make the tautology clear: illustrating a point, not proving a thing since clearly depends on if you assume displacement, and then degree of non-overlap etc.

## Generalized Experiments

We focused first on the range of impacts that MPAs can have on mature
biomass across the range of variables explored. To illustrate, we pulled
out two example simulations, one where MPAs provide relatively large
benefits across all three species, and a second where MPAs actually
produced substantial losses in some species. In the "best case"
scenario, each of the three species shares a common habitat
distribution, and so protection of the central habitat of one species
results in protection for all species. In the "worst case" scenario,
each of the species have markedly different centroids. An MPA designed
primarily around the tuna and marlin populations then ends up displacing
effort into the shark habitat, producing benefits for the first two
species but driving the shark population down.

@@

knitr::include_graphics(file.path(results_path,"fig_info.pdf"))
```

the point is not once answer

Looking at the full
`r n_distinct(total_experiment_results$placement_id) * n_distinct(total_experiment_results$state_id)`
experiments run, the results were highly variable, with MPAs capable of
producing substantial conservation gains to substantial conservation
losses and everything in between. We do not focus into the individual
effects of particular variables as we found them to be highly dependent
on the exact question asked and the set of data included, and since we
cannot ascribe any actual likelihood to any individual state of nature
simulated here.

@@

\\Big message: even constrained by the same initial depletion and core
biology (e.g. movement rates), a wide range of outcomes are possible for
each species for a given MPA size, dependent on the design strategy,
habitat overlap, migrations, quality of information, etc. Across the
board as MPA size increases the ratio of positive to negative outcomes
increases, until all species achieve their maximum conservation
potential in every iteration at 100% MPA. HOwever, every species
experienced a set of simulations, roughyl xx%, that resulted in net
losses in SSB/SSB0 relative to BAU. Losses were smaller than potential
gaines, but reached meaningful levels in some species, and interestingly
peaked in many cases right before max MPA (reflecting concentration of
fishing in small amounts of habitat in habited by the species).Important
to stress in here that there are no probabilities in this, no way of
knowing which state is more probable than others, just given the range
of things simulated more ways for a bigger positive effect. Also make
clear though, biggish losses (10% or more) possible, but rare.

@@

To address these questions, we ran a more limited "case study" set of
simulations, based on the dynamics of the WCPO tuna fisheries. Rather
than iterating over different movement dynamics and habitat
configurations, we fix habitat layers based on reported CPUE for the
species in question, and xx other things we fix xx.
